White Dwarf Mergers on Adaptive Meshes\\ II. Collisions

### 2016-07-18 papers/wdmerger_II/paper.tex
Write up the collision network dependence

Now we move to a discussion of the effect of including various isotopes
in the nuclear reaction network. In particular we compare the \isoseven, \aproxthirteen,
\aproxnineteen, and \aproxtwentyone\ reaction networks. In all cases we run
the same problem setup, and the unused isotopes (everything but carbon and oxygen)
have their mass fractions in every zone set to a negligibly small number at initialization.

>>>>

Now we move to a discussion of the effect of including various isotopes
in the nuclear reaction network. In particular we compare the \isoseven, \aproxthirteen,
\aproxnineteen, and \aproxtwentyone\ reaction networks. In all cases we run
the same problem setup, and the unused isotopes (everything but carbon and oxygen)
have their mass fractions in every zone set to a negligibly small number at initialization.
\autoref{table:networks} lists the $^{56}$Ni generation and total energy generation (determined
by subtracting the initial energy from the final energy). The energy generated in the latter
three cases agrees remarkably well, with a less than $0.5\%$ difference. \aproxthirteen\
over-generates nickel at the $2\%$ level compared to the networks with more isotopes, a reasonably
small difference. On the other hand, agreement with \isoseven\ is not very strong, as \isoseven\
generates far too much nickel and far too little energy. Nevertheless while the quantitative
agreement is not there, the collision looks qualitatively similar compared to the evolution with
more complicated networks, so while \isoseven\ should not be used for predicting nucleosynthetic
yields in production science runs, it may still be fair for use in proof-of-concept studies.

\input{plots/networks_m_P_0.64_m_S_0.64.tbl}

\isoseven\ also has the peculiar property that after the detonation had passed through the WDs,
the region at the central contact point did not fully stop burning, yielding a quasistatic energy
release that added a couple of million extra timesteps to the simulation (in addition to the few
hundred it normally takes) before we manually terminated the run. The full effect of these additional
timesteps only amounted to a $10^{-4}$ relative change in the nickel production. By examining the behavior
of zones near the collision point, we discovered the reason for this. The zones in question have a composition that
is approximately $98\%$ Ni and $2\%$ He by mass. For this combination, a direct evaluation of the RHS in
comparison to, say, \aproxthirteen, reveals that $dX/dt$ for these isotopes is off by many orders of magnitude.
This is a consequence of the choice inherent to the network to assume that the isotopes in between $^{28}$Si
and $^{56}$Ni are in an effective equilibrium state. So an advective update that yields even a small change
to the abundances in the zone will very strongly knock the zone out of equilibrium with respect to the burn
step following the advective update. This then causes the burning timestep limiter to kick in and make the
timestep very small so that the advective changes do not cause such strong effects on the burning equilibrium.
This effect is an argument against the use of \isoseven\ for explosive, non-hydrostatic burning problems.
